Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
fix algs

    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \MergeLOf{\Lens_1}{\Lens_2}, \Regex}$
    & =
    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + $\\
    &
    & $(1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \ProbabilityAlt))+$\\
    &
    & $(0, -\Log_2(\ProbabilityAlt)*\ProbabilityAlt -\Log_2(1 - \ProbabilityAlt)*(1-\ProbabilityAlt)]$\\

@@

The cost of a lens between two SREs $c(\Lens, \Regex, \RegexAlt) =
\MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the maximum sum of
recovering the left format from the right, and the right from the left. We have
proven theorems demonstrating the calculated entropy corresponds to the actual
conditional entropy to recover the data.

@@

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a
high level of abstraction. The
input regular expressions are first converted into stochastic regular
expressions with \ToStochastic. Once that conversion has taken place,
the $\RXSearch$ algorithm executes.
This algorithm itself is defined via
five functions:
(1) $\PCF{\RXSearch.Create}$,
(2) $\PCF{\RXSearch.Continue}$,
(3) $\PCF{\RXSearch.Propose}$,
(4) $\PCF{\RXSearch.Enqueue}$, and
(5) $\PCF{\RXSearch.Best}$.
First, $\PCF{\RXSearch.Create}$ creates a priority queue ($rxs$)
containing a single element:  the initial pair of stochastic regular
expressions.  Next, 
$\RXSearch$ enters a loop that searches for likely lenses.  The loop
terminates when the algorithm believes it is unlikely
to find a better lens than the best one it has found so far (a
termination condition defined by $\PCF{\RXSearch.Continue}$).
Within each iteration of the loop, it:
\begin{itemize}
\item generates the next class ($S$, $T$) of lenses to
search ($\PCF{\RXSearch.Propose}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($lco$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification, and
\item adds this best lens to the set of possibilities
under consideration $\PCF{\RXSearch.Add}$
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($\PCF{\RXSearch.Best}$).  Each element of this algorithm will be explained in
further depth in the following subsections.

\dpw{Note: I added overbars to regexps in Algorithm 1 below ... double check me
for correctness.}
\begin{algorithm}
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\BRegex,\BRegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\BRegex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\BRegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Create}{\Regex,\RegexAlt}$
    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \State $\RXSearchState \gets \Call{\RXSearch.Add}{\RXSearchState,\LCO}$
    \EndWhile
    \State \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\subsection{\PCF{\RXSearch.Propose}}
The first synthesizer is one that searches through regular expression pairs.
This synthesizer, \RXSearch, proposes regular expressions, semantically but not
syntactically equivalent to the originals.  In particular, this synthesizer
performs the following rewrites:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

@@

\RXSearch proposes rewrites incrementally, where it proposes regular expression
pairs with the smallest number of rewrites first. When there are no more
proposed regular expressions derived from $n$ rewrites, \RXSearch proposes
regular expressions derived from $n+1$ rewrites.

\RXSearch terminates when the sum of the number of regular expressions of the
smallest unproposed size with the distance of those regular expressions from the
originals exceeds the cost of the current best lens. At this point, \RXSearch
stops proposing regular expression pairs, and instead returns to the user the
best lens found thus far. Note this means that if \RXSearch finds a bijective
lens, which has zero cost it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
for stochastic regular expressions.  Of all the star-semiring equations
required to relate two \SDNFREabbrevs, only those relating to Kleene star
(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs, other axioms, like associativity, commutativity
or distributivity, are unneeded.  This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed
by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
same ``class'' (\IE, any lenses that have the same type when the
Kleene star axioms are unused).  This property, and the resulting
efficiency of $\GreedySynth$ is key to the effectiveness of our
core algorithms.

@@

Intuitively, stochastic DNF regular expressions are essentially stochastic
regular expressions with all concatenations fully distributed over all
disjunctions. As such, the language of a stochastic DNF regular expression is a
union of its subcomponents, the language of a stochastic sequence is the
concatenation of its subcomponents, and the language of a stochastic atom is the
iteration of its subcomponent.

@@

  \AtomToDNF{} \OfType
  \ArrowTypeOf{\AtomType{}}{\DNFRegexType{}}\\
  $\AtomToDNFOf{\Atom} = \DNFOf{(\SequenceOf{\EmptyString \SeqSep \Atom \SeqSep
      \EmptyString},1)}$
  \caption{DNF Regular Expression Functions}
  \label{fig:dnf-regex-functions}
\end{figure}

@@

\[
  \begin{array}{rl}
    & \REntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
    & \hspace*{1.21em}\Given
      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
      \ldots\DNFLSep
      (i_p,j_p,\SSQLens_p)}
      ,\ListOf{c_1;\ldots;c_q}
      ,\ListOf{d_1;\ldots;d_r})\\
    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
    & \text{where } H_j = \frac{\Sigma_{k\SuchThat j_k = j}
      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j}}{\SizeOf{\SetOf{k\SuchThat
      j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}}
  \end{array}
\]

>>>>

    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \MergeLOf{\Lens_1}{\Lens_2}, \Regex}$
    & =
    & $(0,\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}+\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2}+1)$\\

@@

The cost of a lens between two SREs $cost(\Lens, \Regex, \RegexAlt) =
\MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the maximum sum of
recovering the left format from the right, and the right from the left. We have
proven theorems demonstrating the calculated entropy corresponds to the actual
conditional entropy to recover the data.

@@

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
level of abstraction. The input regular expressions are first converted into
stochastic regular expressions with \ToStochastic. Once that conversion has
taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
$\PCF{\PQ.Create}$ creates a priority queue ($pq$) containing a single element:
the initial pair of stochastic regular expressions. Next, $\PCF{SynthSymLens}$
enters a loop that searches for likely lenses. The loop terminates when the
algorithm believes it is unlikely to find a better lens than the best one it has
found so far (a termination condition defined by $\PCF{\Continue}$). Within each
iteration of the loop, it:
\begin{itemize}
\item generates the next class ($S$, $T$) of lenses to
search ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost, and
\item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue with $\PCF{\PQ.Push}$
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($\PCF{\RXSearch.Best}$).  Each element of this algorithm will be explained in
further depth in the following subsections.

\begin{algorithm}
  \caption{\SynthSymLens}
  \label{alg:synth-sym-lens}
  \begin{algorithmic}[1]
    \Function{SynthSymLens}{$\BRegex,\BRegexAlt,\Examples$}
    \State $\Regex \gets \Call{\ToStochastic}{\BRegex}$
    \State $\RegexAlt \gets \Call{\ToStochastic}{\BRegexAlt}$
    \State $\RXSearchState \gets \Call{\PQ.Create}{\Regex,\RegexAlt}$
    \State $\Best \gets \None$
    \While{$\Call{\Continue}{\RXSearchState,\Best}$}
    \State $(\Regex,\RegexAlt) \gets \Call{\PQ.Pop}{\RXSearchState}$
    \State $\LC \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
    \If{$\Call{Cost}{\LC} < \Call{Cost}{\Best}$}
    \State $\Best \gets \LC$
    \EndIf
    \State $\Call{\PQ.Push}{\RXSearchState,\Expand(\Regex,\RegexAlt)}$
    \EndWhile
    \State \ReturnVal{best}
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\subsection{Search Through Types}
The first synthesizer is one that searches through regular expression pairs.
This synthesizer, comprising \Expand, \PQ, and \RXSearch, proposes regular
expressions, semantically but not syntactically equivalent to the originals. In
particular, this synthesizer performs the following rewrites:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

@@

Our priority queue proposes stochastic regular expressions generated from fewer
rewrites first. Only when there are no more proposed regular expressions derived
from $n$ rewrites will \PCF{\PQ.Pop} propose regular expressions derived from
$n+1$ rewrites.

The procedure \PCF{\Continue} terminates when the sum of the number of regular expressions of the
smallest unproposed size with the distance of those regular expressions from the
originals exceeds the cost of the current best lens. At this point, \RXSearch
stops proposing regular expression pairs, and instead returns to the user the
best lens found thus far. Note this means that if \RXSearch finds a bijective
lens, which has zero cost it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
for stochastic regular expressions.  Of all the star-semiring equations
required to relate two \SDNFREabbrevs, only those relating to Kleene star
(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs, other axioms, like associativity, commutativity
or distributivity, are unneeded.  This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed
by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
same ``class'' (\IE, any lenses that have the same type when the
Kleene star axioms are unused).  This property, and the resulting
efficiency of $\GreedySynth$ is key to the effectiveness of our
core algorithms.

@@

Intuitively, stochastic DNF regular expressions are essentially stochastic
regular expressions with all concatenations fully distributed over all
disjunctions. As such, the language of a stochastic DNF regular expression is a
union of its subcomponents, the language of a stochastic sequence is the
concatenation of its subcomponents, and the language of a stochastic atom is the
iteration of its subcomponent. For
$\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}$
to be a valid stochastic DNF regular expression, the probabilities must sum to
one ($\sum_{i=0}^n\Probability_i = 1$).

@@

  \AtomToDNF{} \OfType
  \ArrowTypeOf{\AtomType{}}{\DNFRegexType{}}\\
  $\AtomToDNFOf{\Atom} = \DNFOf{(\SequenceOf{\EmptyString \SeqSep \Atom \SeqSep
      \EmptyString},1)}$
  \caption{Stochastic DNF Regular Expression Functions}
  \label{fig:dnf-regex-functions}
\end{figure}

@@

\[
  \begin{array}{rl}
    & \REntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
    & \hspace*{1.21em}\Given
      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
      \ldots\DNFLSep
      (i_p,j_p,\SSQLens_p)}
      ,\ListOf{c_1;\ldots;c_q}
      ,\ListOf{d_1;\ldots;d_r})\\
    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
    & \text{where } H_j = (0,\Sigma_{\SetOf{k\SuchThat j_k = j}}
      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j})
  \end{array}
\]
